Modeling the Evolution of Mimicry

### 2016-02-24 docs/paper/Paper.tex
Revert "More formtting"
This reverts commit a11fb7808ed8b619a9306377c9ad2917659f5e55.

%Reformat current date and time
\usepackage{datetime}

%For hyper referencing
\usepackage[pdfborder={text},colorlinks={true},linkcolor={black},citecolor={black},urlcolor={black}]{hyperref}
%In case you use the package hyperref to create a PDF, the links to tables or figures will point to the caption of the table or figure, which is always below the table or figure itself. Therefore the table or figure will not be visible, if it is above the pointer and one has to scroll up in order to see it. If you want the link point to the top of the image you can use the package hypcap with:
\usepackage[all]{hypcap}

@@

Section \ref{section:mimicry} discusses an inspiration from the field of evolutionary biology which is the primary purpose of building the model of this thesis. Mimicry is an evolutionary process, which not only provides us with insight into the process of evolution as it happens in nature but also is fascinating to study and learn how nature adapts itself in complicated situation for the survival of its species. In addition, this chapter discusses the different kinds of mimicry such as Batesian and Mullerian, while including how the concept of mimicry ring helps us explain each of these cases.

@@

This chapter contains a discussion on the field of Evolutionary Computing and Artificial Life. Both of these field of computer science are inspired by the process of evolution as it happens in nature. The model presented in this thesis is also based on an evolutionary process existing in nature. This thesis is part of the field of Artificial Life, motivated by the work of Echo \cite{holland1996} and other such simulated Complex Adaptive Systems. In addition to the model discussed in chapter \ref{section:model}, this thesis also put some effort in finding application for the evolutionary process of mimicry (chapter \ref{section:application}). Similar to evolutionary computation, we intend to apply the deceptive process to solve a problem in computer science.

@@

\subsubsection{Complex Adaptive System}
\label{subsec:complex-adaptive-system}
\gls{cas} are special cases of complex systems. They are complex as they are diverse and made up of multiple interconnected elements and adaptive as they have the capacity to change and learn from experience. Examples of complex adaptive systems include the stock market, social insect and ant colonies, the biosphere and the ecosystem, the brain and the immune system, the cell and the developing embryo, manufacturing businesses and any human social group-based endeavor in a cultural and social system such as political parties or communities. This includes some large-scale online systems, such as collaborative tagging or social bookmarking systems.

@@

\subsubsection{Echo}
\label{subsec:echo}
Echo is a class of simulation model for complex adaptive system, designed primarily for gedanken experiments rather than precise simulations. Echo provides a population of evolving, reproducing agents distributed over a geography, with different inputs of renewable resources at various sites. Each agent has simple capabilities - offense, defense, trading and mate selection - defined by a set of chromosomes. Even though these capabilities are simple and defined simply, they provide a rich set of variations illustrating the four kernel properties of complex adaptive systems described in section \ref{subsec:complex-adaptive-system}.

@@

\subsection{Conclusion}
This chapter discusses evolutionary computation, the field that takes concepts from evolutionary biology to solve and optimize problems in compute science. It also discusses the interdisciplinary field Artificial life where scientist discuss diverse knowledge from biology, economics, physics and use computer science to simulate and analyze life as it is in nature. The discussion of this chapter is expected to provide a base upon which the following chapters will begin construction of the core idea of this thesis.

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=1]{../tex/images/Batesplate_ArM}
	\caption[Plate from Bates (1862) illustrating Batesian mimicry]{Plate from Bates (1862) illustrating Batesian mimicry between Dismorphia species (top row, third row) and various Ithomiini (Nymphalidae) (second row, bottom row). \cite{bates1862}}
	\label{fig:batesian-butterfly}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.75]{../tex/images/BatesMimButter}
	\caption[A very well-known example of mimicry, the viceroy butterfly]{A very well-known example of mimicry is the viceroy butterfly (top) which has a pattern very similar to the unpalatable monarch butterfly (bottom). It was considered for a long time that this is an ideal example of Batesian mimicry. But recently Brower discovered that the viceroy is actually just as noxious as the monarch, making this a case of Mullerian mimicry. \cite{brower1991} Image source: \href{http://en.wikipedia.org/wiki/Mullerian_mimicry}{Wikipedia}}
	\label{fig:mullerian-butterfly}
\end{figure}

\subsection{Evolutionary Dynamics of Mimicry}
\label{sec:evolutionary-dynamics-of-mimicry}
The dynamics of mimicry has been investigated by Turner \cite{turner1988}, where he states that the evolution of mimicry can be explained best by the process of punctuated equilibrium instead of phyletic gradualism.

@@

\subsubsection{Models by Franks and Noble}
\label{subsec:models-by-frank-and-noble}
%Explain more about the first model by Franks and Noble
The \textbf{first} model by Franks and Noble \cite{franks2002} is where \textsl{``multiple species are preyed upon by a single abstract predator; the appearance of each prey species can evolve but their palatability is fixed"}. Each individual had a single gene: a value representing their external appearance or phenotype. The phenotypes are constrained to a ring of values from 1-20 (where 20 and 1 are neighbors). The distance of one phenotype from another represents their levels of similarity.

@@

\begin{table}[H]
\centering
\begin{tabular}{| p{2.2cm} | >{\centering} p{3cm} | p{7.5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		ISize & 6 & Number of cell in single dimension of the 3-D toroidal cube\\ \hline
		World Size & 20 & Size of a single dimension of the 3-D toroidal cube\\ \hline
		Cell Size & \( World Size / ISize \) & Size of each cell\\ \hline
		Total Number of Cells & \( ISize^3  = 216\) & Total number of cells in the environment\\ 
	\hline
\end{tabular}
\caption{Parameters to control the environment.}
\label{tab:environment-control-parameters}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{cc}
		\multicolumn{2}{c}{ \includegraphics[scale=0.60]{../tex/images/cells-front} } \\
		\includegraphics[scale=0.50]{../tex/images/cells-top} & \includegraphics[scale=0.50]{../tex/images/cells-side} \\
\end{tabular}
\caption[Three dimensional representation of the environment]{Three dimensional representation of the environment divided in cells. Presence of different species of agents inside.}
\label{tab:3-d-environment-images}
\end{table}

@@

\begin{table}[H]
\centering
\begin{tabular}{| p{2.2cm} | >{\centering} p{1.3cm} | p{9cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Force Factor & 40 & A unit vector is multiplied by this amount before being added to the force vector.\\ \hline
		Differential Time Step (DT) & 0.01 & The time step used for first-order integration of the motion equations.\\ \hline
		Friction & 5 & The friction constant used for motion calculations.\\ \hline
		Work Factor & 1 & \( Work done = WF * force * distance \), where \(WF\) is this constant.\\
	\hline
\end{tabular}
\caption{Parameters to control mobility of agents.}
\label{tab:mobility-control-parameters}
\end{table}

@@

\begin{table}[H]
	\centering
	\begin{tabular}{| l | c | c | c | c | c | c | c | c |}
	  \hline
	  Current Pattern & 111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\ \hline
	  New state of center cell & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 \\
	  \hline
	\end{tabular}
	\caption{Cellular Automata rule}
	\label{tab:cellular-automata-rule}
\end{table}

@@

\subsubsection{Species diversity}
\label{subsec:species-diversity}
Franks and Noble have used different models to diversify species. In \cite{franks2002} they have used linear difference in number ``constrained to a 'ring' of values  from 1-20 (where 20 and 1 are neighbors)" to distinguish species diversity. Here the distance of one phenotype from another represents their level of similarity.

@@

With CA based pattern representation, population of prey species with a specific CA pattern can be grouped as one single species. And by restricting inter species reproduction we can control the diversity of patterns. But there is mutation applied when similar species mate with each other, so new species do born out of generations of existing species. So that is why we have two separate mutation rate while reproducing prey species. One being the ``Pattern Mutation Rate" (default values are mentioned in table \ref{tab:prey-control-parameters}) with which we control mutation of the first 8 bits of the genome while the ``Genome Mutation Rate" is used to control mutation of rest of the 9 bit genome (table \ref{tab:prey-genome}). Similar efforts of multiple mutation rate at varying location has been used in developing Echo as mentioned in section \ref{subsec:echo} \cite{hraber1997}.

\subsubsection{Genome}
The Genome of the prey species consists of 17 bits as presented in table \ref{tab:prey-genome}. The first eight bits represent the rule, which is used to generate Cellular Automata pattern. The next two bits are used to represent palatability of the organism explained detail in section \ref{sec:genetic-palatability-representation}. The next six bits is the magnitude of the force with which mobility of the organism is calculated. The 17th bit is used to evaluate reproduction capability of the organism, depending on which the prey species is either capable or not capable to participate in reproduction.

\begin{table}[H]
\centering
\begin{tabular}{|c|c|c|c|}
	\hline
		\textbf{Pattern(8)} & \textbf{Palatability(2)} & \textbf{Mobility(6)} & \textbf{Reproduction(1)} \\ \hline
		10101101					 	& 							01		 		 & 			110001					&					1						 		 \\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 17 bit prey genome.}
\label{tab:prey-genome}
\end{table}

\subsubsection{Reflection of punctuated equilibrium}
\label{subsec:reflection-of-punctuated-equilibrium}
Punctuated equilibrium has been defined in section \ref{sec:evolutionary-dynamics-of-mimicry} as the evolutionary process that is more inclined to cladogenesis instead of gradualism. Also Turner \cite{turner1988} emphasizes on punctuated equilibrium to describe the evolution of mimicry instead of phyletic gradualism. The design of the model under discussion also follows Turner's explanation in terms of evolving mimicry. As it can be observed, new CA patterns evolve from existing ones in prey population just by a single mutation in the pattern gene. Mimics do not follow a gradual process of evolution to look close to models but rather the change happens randomly through a single step mutation. The mutations that are favored, helps the mimics to survive while the unfavored ones fail to persist. It can be observed later in table \ref{tab:diff-in-pattern} how CA patterns of prey species can have vastly different configuration for a unit change in their representative gene.

\subsubsection{Genetic representation of palatability}
\label{sec:genetic-palatability-representation}
The palatability of each prey species is fixed and has been represented with 2 bits of the genome giving it a range of 0 to 3 with four levels of palatability. The combinations are as follows:

\begin{table}[H]
	\centering
	\begin{tabular}{|c|c|}
		\hline
			\textbf{Gene (Index 8 to 9)} &	\textbf{Palatable} \\ \hline
			00									& True 			\\ \hline
			01									& True 			\\ \hline
			10									& False 		\\ \hline
			11									& False 		\\
		\hline
	\end{tabular}
	\caption{Genetic representation of palatability}
	\label{tab:genetic-representation-palatability}
\end{table}

@@

\begin{algorithm}[H]
	\caption{Algorithm for reproduction of the Prey species}
	\label{algo:algorithm-reproduction-prey}
	\begin{algorithmic}
		\FOR{each step in time}
			\STATE $capableToReproduce \gets$ true or false depending on the 17th bit of the genome, and maturity on reproduction age.
			\IF {$capableToReproduce == true$ and $currentCellPopulation > 1$ } 
				\STATE $anotherPrey \gets$ Select random prey from same cell.
				\IF {$anotherPrey$ is alive and $capableToReproduce$ and of similar $pattern$ and $palatability$}
					\STATE Perform genetic cross over and mutation to create new genome.
					\STATE Create new prey with new genome, and release into environment.
					\STATE Record reproduction time for both prey.
				\ENDIF
			\ENDIF
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

\begin{table}[H]
\centering
\begin{tabular}{| p{2.2cm} | >{\centering} p{1.5cm} | p{9cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Prey Size & 2 to 5 & Size of the prey species in the 3D FormAL  environment.\\ \hline
		Reproduction age limit & 100 & Minimum number of iterations or time steps a prey species need to be present in the simulation to get reproduction capability\\ \hline
		Reproduction interval & 1000 & Number of iterations a prey need to wait before reproducing again.\\ \hline
		Pattern Mutation Rate & 0.05 & Rate of Mutation of the pattern genome.\\ \hline
		Genome Mutation Rate & 0.5 & Rate of mutation of the rest of the genome excluding the pattern gene.\\ \hline
		Demise Age & 2000 & Age at which the prey species will be removed from the environment.\\
	\hline
\end{tabular}
\caption{Parameters to control prey population and visibility.}
\label{tab:prey-control-parameters}
\end{table}

@@

\begin{algorithm}
	\caption{Algorithm for attacking Prey species}
	\label{algo:algorithm-attack-prey}
	\begin{algorithmic}
		\FOR{each step in time}
			\IF {Predator has reached Minimum attack age}
				\STATE $preyToAttack \gets$ select random prey species from the same cell.
				\IF {$currentMemorySize < MinimumMemorySize$}
					\STATE $attemptToKill(preyToAttack);$ \COMMENT {Function defined in algorithm \ref{algo:algorithm-attemptToKill}}
				\ELSE
					\STATE recognize $preyToAttack$ pattern from memory.
					\IF {prey is not recognized or prey is palatable}
						\STATE $attemptToKill(preyToAttack)$
					\ELSE 
						\STATE Prey is recognized as unpalatable, so release it back into environment.
					\ENDIF
				\ENDIF
			\ENDIF
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}
	\caption{$attemptToKill(preyToAttack)$}
	\label{algo:algorithm-attemptToKill}
	\begin{algorithmic}
		\STATE Kill $preyToAttack$.
		\IF {Memory size $< MaxMemorySize$}
			\STATE Associate $preyToAttack$ pattern and palatability and store into memory.
			\STATE Recalculate Hopfield Memory Weights using Hebbian Learning outer product rule.
		\ENDIF
	\end{algorithmic}	
\end{algorithm}

@@

\begin{table}[H]
\centering
\begin{tabular}{|c|c|}
	\hline
		\textbf{Mobility(4)} & \textbf{Reproduction(1)} \\ \hline
				 1101					   &					1						 		\\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 5 bit predator genome.}
\label{tab:predator-genome}
\end{table}

@@

\begin{table}[h!]
\centering
\begin{tabular}{| p{2.2cm} | >{\centering} p{2.2cm} | p{8cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Minimum Memory Size & 2 to 6 & Minimum number of patterns stored in memory before predators start making intelligent decisions.\\ \hline
		Maximum Memory Size & 10 & Maximum number of patterns to be stored in memory. Limited to reduce processing time. \\ \hline 
		Hopfield Maximum Iterations & 20 & Maximum number of iterations for Hopfield Network to recognize a pattern. Usually the network reaches a steady state before that. But this restriction is to avoid infinite loop in case the network never reaches a steady state. \\ \hline
		Attack Age & 500 & minimum age a predator needs to reach to be able to attack prey species.  \\ \hline
		Attack Interval & 100 & Interval of time which needs to pass before a predators attacks its next prey. \\ \hline
		Genome Mutation rate & 0.3 & Mutation rate for the 5 bit genome of the predators representing their mobility and pattern recognition capability. \\ \hline
		Reproduction Age Limit & 500 & Minimum age a predator needs to reach before engaging in reproduction.\\ \hline
		Reproduction Interval & 1000 to 3000 & Interval of time a predator needs to pass between two reproduction process.\\ \hline
		Demise Age & 2000 to 7000 & Age at which a predator is considered as dead.\\
	\hline
\end{tabular}
\caption{Parameters to control predator population and pattern recognition capability.}
\label{tab:predator-control-parameters}
\end{table}

\subsection{Conclusion}
This model has been designed from a practical point of view; to come up with efficient results and achieve the main objective, \textit{evolution of mimicry}. Creation and transformation of different mimicry ring and also the dynamics of it has been integrated to achieve interesting results. This model can also be considered as a complex adaptive system similar to Holland's work on Echo \cite{holland1996}. The seven basics of a complex adaptive system which are: Aggregation, Tagging, Nonlinearity, Flow, Diversity, Internal Models and Building blocks \cite{holland1996} are present in this model. Individual components of this model such as the different types of agents and their properties can be considered as \textit{building blocks}. Each prey species are \textit{tagged} with individual pattern and palatability with which predators recognize them. We are providing different properties, behaviors and goals to the agents but setting them free in the environment to observe their \textit{aggregate} behavior, resulting in \textit{non-linear} or unpredictable outcome. The model has its \textit{flow} as it progresses in time. Also there is \textit{diversity} of prey species in the environment (\ref{subsec:species-diversity}).

@@

\begin{table}[H]
\centering
\begin{tabular}{| p{2cm} | >{\centering} p{2.2cm} | p{8cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Mimicry Ring Hamming Distance & 10 \% of the Pattern Size & If the Hamming distance between the pattern of the model and the mimic is 10 \% the size of the pattern then they are considered within the same ring.\\ \hline
		Number of Rings to report & 8 & This value is the number of most populous rings that are included in the report.\\
	\hline
\end{tabular}
\caption{Parameters to mimicry ring report.}
\label{tab:ring-report-control-parameters}
\end{table}

>>>>

%For hyper referencing
\usepackage[colorlinks={true},linkcolor={black},citecolor={black},urlcolor={black}]{hyperref}
%In case you use the package hyperref to create a PDF, the links to tables or figures will point to the caption of the table or figure, which is always below the table or figure itself. Therefore the table or figure will not be visible, if it is above the pointer and one has to scroll up in order to see it. If you want the link point to the top of the image you can use the package hypcap with:
\usepackage[all]{hypcap}

@@

Section \ref{section:mimicry} discusses an inspiration from the field of evolutionary biology which is the primary purpose of building the model of this thesis. Mimicry is an evolutionary process, which not only provides us with insight into the process of evolution as it happens in nature but also is fascinating to study and learn how nature adapts itself in complicated situation for the survival of its species. In addition, this section discusses the different kinds of mimicry such as Batesian and Mullerian, while including how the concept of mimicry ring helps us explain each of these cases.

@@

This section contains a discussion on the field of Evolutionary Computing and Artificial Life. Both of these field of computer science are inspired by the process of evolution as it happens in nature. The model presented in this thesis is also based on an evolutionary process existing in nature. This thesis is part of the field of Artificial Life, motivated by the work of Echo \cite{holland1996} and other such simulated Complex Adaptive Systems. In addition to the model discussed in section \ref{section:model}, this thesis also put some effort in finding application for the evolutionary process of mimicry (section \ref{section:application}). Similar to evolutionary computation, we intend to apply the deceptive process to solve a problem in computer science.

@@

\subsubsection{Complex Adaptive System}
\label{subsubsec:complex-adaptive-system}
\gls{cas} are special cases of complex systems. They are complex as they are diverse and made up of multiple interconnected elements and adaptive as they have the capacity to change and learn from experience. Examples of complex adaptive systems include the stock market, social insect and ant colonies, the biosphere and the ecosystem, the brain and the immune system, the cell and the developing embryo, manufacturing businesses and any human social group-based endeavor in a cultural and social system such as political parties or communities. This includes some large-scale online systems, such as collaborative tagging or social bookmarking systems.

@@

\subsubsection{Echo}
\label{subsubsec:echo}
Echo is a class of simulation model for complex adaptive system, designed primarily for gedanken experiments rather than precise simulations. Echo provides a population of evolving, reproducing agents distributed over a geography, with different inputs of renewable resources at various sites. Each agent has simple capabilities - offense, defense, trading and mate selection - defined by a set of chromosomes. Even though these capabilities are simple and defined simply, they provide a rich set of variations illustrating the four kernel properties of complex adaptive systems described in section \ref{subsubsec:complex-adaptive-system}.

@@

\subsection{Conclusion}
This section discusses evolutionary computation, the field that takes concepts from evolutionary biology to solve and optimize problems in compute science. It also discusses the interdisciplinary field Artificial life where scientist discuss diverse knowledge from biology, economics, physics and use computer science to simulate and analyze life as it is in nature. The discussion of this section is expected to provide a base upon which the following sections will begin construction of the core idea of this thesis.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/Batesplate_ArM}
	\caption[Plate from Bates (1862) illustrating Batesian mimicry]{Plate from Bates (1862) illustrating Batesian mimicry between Dismorphia species (top row, third row) and various Ithomiini (Nymphalidae) (second row, bottom row). \cite{bates1862}}
	\label{fig:batesian-butterfly}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.75]{../tex/images/BatesMimButter}
	\caption[A very well-known example of mimicry, the viceroy butterfly]{A very well-known example of mimicry is the viceroy butterfly (top) which has a pattern very similar to the unpalatable monarch butterfly (bottom). It was considered for a long time that this is an ideal example of Batesian mimicry. But recently Brower discovered that the viceroy is actually just as noxious as the monarch, making this a case of Mullerian mimicry. \cite{brower1991} Image source: \href{http://en.wikipedia.org/wiki/Mullerian_mimicry}{Wikipedia}}
	\label{fig:mullerian-butterfly}
\end{figure}

\subsection{Evolutionary Dynamics of Mimicry}
\label{subsec:evolutionary-dynamics-of-mimicry}
The dynamics of mimicry has been investigated by Turner \cite{turner1988}, where he states that the evolution of mimicry can be explained best by the process of punctuated equilibrium instead of phyletic gradualism.

@@

\subsubsection{Models by Franks and Noble}
\label{subsubsec:models-by-frank-and-noble}
%Explain more about the first model by Franks and Noble
The \textbf{first} model by Franks and Noble \cite{franks2002} is where \textsl{``multiple species are preyed upon by a single abstract predator; the appearance of each prey species can evolve but their palatability is fixed"}. Each individual had a single gene: a value representing their external appearance or phenotype. The phenotypes are constrained to a ring of values from 1-20 (where 20 and 1 are neighbors). The distance of one phenotype from another represents their levels of similarity.

@@

\begin{table}[h]
\centering
\begin{tabular}{| p{1.5cm} | >{\centering} p{1.6cm} | p{4cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		ISize & 6 & Number of cell in single dimension of the 3-D toroidal cube\\ \hline
		World Size & 20 & Size of a single dimension of the 3-D toroidal cube\\ \hline
		Cell Size & \( World Size\) / \(ISize \) & Size of each cell\\ \hline
		Total Number of Cells & \( ISize^3  = 216\) & Total number of cells in the environment\\ 
	\hline
\end{tabular}
\caption{Parameters to control the environment.}
\label{tab:environment-control-parameters}
\end{table}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.5\textwidth]{../tex/images/cells-front}
	\caption[Three dimensional representation of the environment]{Three dimensional representation of the environment divided in cells. Presence of different species of agents inside.}
	\label{tab:3-d-environment-images}
\end{figure}

@@

\begin{table}[H]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{1.5cm} | >{\centering} p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Force Factor & 40 & A unit vector is multiplied by this amount before being added to the force vector.\\ \hline
		Differential Time Step (DT) & 0.01 & The time step used for first-order integration of the motion equations.\\ \hline
		Friction & 5 & The friction constant used for motion calculations.\\ \hline
		Work Factor & 1 & \( Work done = WF * force * distance \), where \(WF\) is this constant.\\
	\hline
\end{tabular}
\caption{Parameters to control mobility of agents.}
\label{tab:mobility-control-parameters}
\end{table}

@@

\begin{table}[H]
	\centering
	\setlength\tabcolsep{2pt}
	\begin{tabular}{| p{2.2cm} | c | c | c | c | c | c | c | c |}
	  \hline
	  Current Pattern & 111 & 110 & 101 & 100 & 011 & 010 & 001 & 000 \\ \hline
	  New state of center cell & 0 & 0 & 0 & 1 & 1 & 1 & 1 & 0 \\
	  \hline
	\end{tabular}
	\caption{Cellular Automata rule}
	\label{tab:cellular-automata-rule}
\end{table}

@@

\subsubsection{Species diversity}
\label{subsubsec:species-diversity}
Franks and Noble have used different models to diversify species. In \cite{franks2002} they have used linear difference in number ``constrained to a 'ring' of values  from 1-20 (where 20 and 1 are neighbors)" to distinguish species diversity. Here the distance of one phenotype from another represents their level of similarity.

@@

With CA based pattern representation, population of prey species with a specific CA pattern can be grouped as one single species. And by restricting inter species reproduction we can control the diversity of patterns. But there is mutation applied when similar species mate with each other, so new species do born out of generations of existing species. So that is why we have two separate mutation rate while reproducing prey species. One being the ``Pattern Mutation Rate" (default values are mentioned in table \ref{tab:prey-control-parameters}) with which we control mutation of the first 8 bits of the genome while the ``Genome Mutation Rate" is used to control mutation of rest of the 9 bit genome (table \ref{tab:prey-genome}). Similar efforts of multiple mutation rate at varying location has been used in developing Echo as mentioned in section \ref{subsubsec:echo} \cite{hraber1997}.

\subsubsection{Genome}
The Genome of the prey species consists of 17 bits as presented in table \ref{tab:prey-genome}. The first eight bits represent the rule, which is used to generate Cellular Automata pattern. The next two bits are used to represent palatability of the organism explained detail in section \ref{subsubsec:genetic-palatability-representation}. The next six bits is the magnitude of the force with which mobility of the organism is calculated. The 17th bit is used to evaluate reproduction capability of the organism, depending on which the prey species is either capable or not capable to participate in reproduction.

\begin{table}[H]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{|c|c|c|c|}
	\hline
		\textbf{Pattern(8)} & \textbf{Palatability(2)} & \textbf{Mobility(6)} & \textbf{Reproduction(1)} \\ \hline
		10101101					 	& 							01		 		 & 			110001					&					1						 		 \\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 17 bit prey genome.}
\label{tab:prey-genome}
\end{table}

\subsubsection{Reflection of punctuated equilibrium}
\label{subsubsec:reflection-of-punctuated-equilibrium}
Punctuated equilibrium has been defined in section \ref{subsec:evolutionary-dynamics-of-mimicry} as the evolutionary process that is more inclined to cladogenesis instead of gradualism. Also Turner \cite{turner1988} emphasizes on punctuated equilibrium to describe the evolution of mimicry instead of phyletic gradualism. The design of the model under discussion also follows Turner's explanation in terms of evolving mimicry. As it can be observed, new CA patterns evolve from existing ones in prey population just by a single mutation in the pattern gene. Mimics do not follow a gradual process of evolution to look close to models but rather the change happens randomly through a single step mutation. The mutations that are favored, helps the mimics to survive while the unfavored ones fail to persist. It can be observed later in table \ref{tab:diff-in-pattern} how CA patterns of prey species can have vastly different configuration for a unit change in their representative gene.

\subsubsection{Genetic representation of palatability}
\label{subsubsec:genetic-palatability-representation}
The palatability of each prey species is fixed and has been represented with 2 bits of the genome giving it a range of 0 to 3 with four levels of palatability. The combinations are as follows:

\begin{table}[h]
	\centering
	\begin{tabular}{|c|c|}
		\hline
			\textbf{Gene (Index 8 to 9)} &	\textbf{Palatable} \\ \hline
			00									& True 			\\ \hline
			01									& True 			\\ \hline
			10									& False 		\\ \hline
			11									& False 		\\
		\hline
	\end{tabular}
	\caption{Genetic representation of palatability}
	\label{tab:genetic-representation-palatability}
\end{table}

@@

\begin{algorithm}[h]
	\caption{Algorithm for reproduction of the Prey species}
	\label{algo:algorithm-reproduction-prey}
	\begin{algorithmic}
		\FOR{each step in time}
			\STATE $capableToReproduce \gets$ true or false depending on the 17th bit of the genome, and maturity on reproduction age.
			\IF {$capableToReproduce == true$ and $currentCellPopulation > 1$ } 
				\STATE $anotherPrey \gets$ Select random prey from same cell.
				\IF {$anotherPrey$ is alive and $capableToReproduce$ and of similar $pattern$ and $palatability$}
					\STATE Perform genetic cross over and mutation to create new genome.
					\STATE Create new prey with new genome, and release into environment.
					\STATE Record reproduction time for both prey.
				\ENDIF
			\ENDIF
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

\begin{table}[h]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Prey Size & 2 to 5 & Size of the prey species in the 3D FormAL  environment.\\ \hline
		Reproduction age limit & 100 & Minimum number of iterations or time steps a prey species need to be present in the simulation to get reproduction capability\\ \hline
		Reproduction interval & 1000 & Number of iterations a prey need to wait before reproducing again.\\ \hline
		Pattern Mutation Rate & 0.05 & Rate of Mutation of the pattern genome.\\ \hline
		Genome Mutation Rate & 0.5 & Rate of mutation of the rest of the genome excluding the pattern gene.\\ \hline
		Demise Age & 2000 & Age at which the prey species will be removed from the environment.\\
	\hline
\end{tabular}
\caption{Parameters to control prey population and visibility.}
\label{tab:prey-control-parameters}
\end{table}

@@

\begin{algorithm}[h]
	\caption{Algorithm for attacking Prey species}
	\label{algo:algorithm-attack-prey}
	\begin{algorithmic}
		\FOR{each step in time}
			\IF {Predator has reached Minimum attack age}
				\STATE $preyToAttack \gets$ select random prey species from the same cell.
				\IF {$currentMemorySize < MinimumMemorySize$}
					\STATE $attemptToKill(preyToAttack);$ \COMMENT {Function defined in algorithm \ref{algo:algorithm-attemptToKill}}
				\ELSE
					\STATE recognize $preyToAttack$ pattern from memory.
					\IF {prey is not recognized or prey is palatable}
						\STATE $attemptToKill(preyToAttack)$
					\ELSE 
						\STATE Prey is recognized as unpalatable, so release it back into environment.
					\ENDIF
				\ENDIF
			\ENDIF
		\ENDFOR
	\end{algorithmic}
\end{algorithm}

\begin{algorithm}[h]
	\caption{$attemptToKill(preyToAttack)$}
	\label{algo:algorithm-attemptToKill}
	\begin{algorithmic}
		\STATE Kill $preyToAttack$.
		\IF {Memory size $< MaxMemorySize$}
			\STATE Associate $preyToAttack$ pattern and palatability and store into memory.
			\STATE Recalculate Hopfield Memory Weights using Hebbian Learning outer product rule.
		\ENDIF
	\end{algorithmic}	
\end{algorithm}

@@

\begin{table}[h]
\centering
\begin{tabular}{|c|c|}
	\hline
		\textbf{Mobility(4)} & \textbf{Reproduction(1)} \\ \hline
				 1101					   &					1						 		\\ \hline
\end{tabular}
\caption{Distribution and purpose of each gene of the 5 bit predator genome.}
\label{tab:predator-genome}
\end{table}

@@

\begin{table}[h!]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | >{\centering} p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Minimum Memory Size & 2 to 6 & Minimum number of patterns stored in memory before predators start making intelligent decisions.\\ \hline
		Maximum Memory Size & 10 & Maximum number of patterns to be stored in memory. Limited to reduce processing time. \\ \hline 
		Hopfield Maximum Iterations & 20 & Maximum number of iterations for Hopfield Network to recognize a pattern. Usually the network reaches a steady state before that. But this restriction is to avoid infinite loop in case the network never reaches a steady state. \\ \hline
		Attack Age & 500 & minimum age a predator needs to reach to be able to attack prey species.  \\ \hline
		Attack Interval & 100 & Interval of time which needs to pass before a predators attacks its next prey. \\ \hline
		Genome Mutation rate & 0.3 & Mutation rate for the 5 bit genome of the predators representing their mobility and pattern recognition capability. \\ \hline
		Reproduction Age Limit & 500 & Minimum age a predator needs to reach before engaging in reproduction.\\ \hline
		Reproduction Interval & 1000 to 3000 & Interval of time a predator needs to pass between two reproduction process.\\ \hline
		Demise Age & 2000 to 7000 & Age at which a predator is considered as dead.\\
	\hline
\end{tabular}
\caption{Parameters to control predator population and pattern recognition capability.}
\label{tab:predator-control-parameters}
\end{table}

\subsection{Conclusion}
This model has been designed from a practical point of view; to come up with efficient results and achieve the main objective, \textit{evolution of mimicry}. Creation and transformation of different mimicry ring and also the dynamics of it has been integrated to achieve interesting results. This model can also be considered as a complex adaptive system similar to Holland's work on Echo \cite{holland1996}. The seven basics of a complex adaptive system which are: Aggregation, Tagging, Nonlinearity, Flow, Diversity, Internal Models and Building blocks \cite{holland1996} are present in this model. Individual components of this model such as the different types of agents and their properties can be considered as \textit{building blocks}. Each prey species are \textit{tagged} with individual pattern and palatability with which predators recognize them. We are providing different properties, behaviors and goals to the agents but setting them free in the environment to observe their \textit{aggregate} behavior, resulting in \textit{non-linear} or unpredictable outcome. The model has its \textit{flow} as it progresses in time. Also there is \textit{diversity} of prey species in the environment (\ref{subsubsec:species-diversity}).

@@

\begin{table}[h]
\centering
\setlength\tabcolsep{2pt}
\begin{tabular}{| p{2cm} | >{\centering} p{1cm} | p{5cm} |}
	\hline
		\textbf{Parameter} & \textbf{Value} & \textbf{Description} \\ \hline
		Mimicry Ring Hamming Distance & 10 \% of the Pattern Size & If the Hamming distance between the pattern of the model and the mimic is 10 \% the size of the pattern then they are considered within the same ring.\\ \hline
		Number of Rings to report & 8 & This value is the number of most populous rings that are included in the report.\\
	\hline
\end{tabular}
\caption{Parameters to mimicry ring report.}
\label{tab:ring-report-control-parameters}
\end{table}
